---
title: 'Fiscal Impact Update'
subtitle: 'Q3 2021 Advanced Estimate'
author: ' Manuel Alcala Kovalski'
date: '`r glue::glue(format(Sys.Date(), "%A, %B %d %Y")," at ", format(strptime(format(Sys.time(), "%H:%M:%S", tz = "America/New_York"), "%H:%M:%S"), "%I:%M %p"))`'
tz: 'America/New_York'
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
mainfont: SourceSansPro
output:
  html_document:
    toc: true # table of content true
    toc_float: true
    toc_depth: 2
    number_sections: false 
    theme: united
    highlight: zenburn 
    hig.retina: 3
    self_contained: yes
    css: [style.css]

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  echo = FALSE,
  fig.align = 'center',
  warning = FALSE,
  message = FALSE,
  cache = FALSE,
  results='hide',
  fig.keep='all'
)

```

```{r}
librarian::shelf(ggplot2, gt, readr)
plots <- read_rds('data/plots')
```

Testing another change again
```{r fiscal_impact}
plots$fiscal_impact %>% print()
```

```{r, include=TRUE, results = 'show'}

revisions_tbl
summary_tbl
```


# NIPA Consistent Purchases {.tabset .tabset-pills}


## Contributions

```{r nipa_purchases}
plots$federal_purchases_contribution
plots$state_purchases_contribution
```



## Levels 

```{r nipa_purchases_levels}
plots$federal_purchases  
plots$state_purchases 

```


# Grants {.tabset .tabset-pills}

## Contributions

```{r grants}
plots$consumption_grants_contribution
plots$investment_grants_contribution

```

## Levels

```{r grants_levels}
plots$consumption_grants
plots$investment_grants
```

# FIM Consistent Purchases 

## Contributions

```{r fim_purchases}
plots$federal_contribution 
plots$state_contribution  
```



# Taxes

## Corporate taxes {.tabset .tabset-pills}

### Contributions

```{r corp_contribution}
plots$federal_corporate_taxes_contribution
plots$state_corporate_taxes_contribution
```

### Levels

```{r corp_lvl}
plots$federal_corporate_taxes
plots$state_corporate_taxes
```

## Non-corporate taxes {.tabset .tabset-pills}


### Contributions

```{r noncorp_cont}
plots$federal_non_corporate_taxes_contribution
plots$state_non_corporate_taxes_contribution
```

### Levels

```{r noncorp_lvl}
plots$federal_non_corporate_taxes
plots$state_non_corporate_taxes

```

# Transfers

## Total 

### Contributions

```{r transfers}
plots$transfers_contribution
plots$federal_transfers_contribution
plots$state_transfers_contribution

```


## Health Outlays {.tabset .tabset-pills}


### Contributions

```{r health}

plots$federal_health_outlays_contribution
plots$state_health_outlays_contribution
```

### Levels

```{r health_lvl}

plots$federal_health_outlays
plots$state_health_outlays
```

### Components

```{r health_comp}
plots$medicaid
plots$medicaid_grants
plots$medicare
```

## Subsidies (non-ARP) {.tabset .tabset-pills}

### Contributions

```{r subsidies}
plots$subsidies_contribution

```

### Levels

```{r subsidies_lvl}

plots$subsidies
```

## Subsidies (ARP) {.tabset .tabset-pills}
ARP PPP and Provider Relief Fund


### Contribution
```{r subsidies-arp-cont}
plots$federal_aid_to_small_businesses_arp_contribution

```

### Levels

```{r subsidies-arp-level}

plots$federal_aid_to_small_businesses_arp
```

## Unemployment Insurance {.tabset .tabset-pills}

### Contributions

```{r ui_cont}
plots$federal_ui_contribution
plots$state_ui_contribution
```

### Levels

```{r ui_lvl}

plots$federal_ui
plots$state_ui
```

## Other aid to vulnerable households {.tabset .tabset-pills}

 Premium Tax Credits, Ratepayer protection, Assistance for older Americans, COBRA, Emergency Assistance
 
### Contribution
```{r vulnerable}
plots$federal_other_vulnerable_arp_contribution
```

### Level

```{r vulnerable_lvl}
plots$federal_other_vulnerable_arp
```

## Rebate Checks (1st and 2nd round){.tabset .tabset-pills}

### Contribution

```{r rebate_cont}
plots$rebate_checks_contribution

```

### Level

```{r rebate_lvl}
plots$rebate_checks

```

## Rebate Checks (ARP) {.tabset .tabset-pills}



### Contribution

```{r rebate_arp_cont}
plots$rebate_checks_arp_contribution

```

### Level

```{r rebate_arp_lvl}
plots$rebate_checks_arp

```

## Direct aid to households {.tabset .tabset-pills}

Includes Child tax credit, EITC,  Childcare for workers,  Dependent  care for families

### Contribution

```{r aid_cont}
plots$federal_other_direct_aid_arp_contribution
```

### Level

```{r aid_lvl}
plots$federal_other_direct_aid_arp
```

## Social Benefits Remainder {.tabset .tabset-pills}



### Contribution
```{r soc_cont}
plots$federal_social_benefits_contribution
plots$state_social_benefits_contribution
```

### Level
```{r soc_lvl}

plots$federal_social_benefits
plots$state_social_benefits
```




# Deflators {.tabset .tabset-pills}

## Delta
```{r}
diff_plot
```

## Current

```{r}
cur_plot
```

## Previous
```{r}
prev_plot
```

<br>
