---
title: 'FIM Update'
author: ' Manu, Georgia, Sarah, and Chase '
date: "`r glue::glue(format(Sys.Date(), '%A, %B %d %Y'), ' at ', format(Sys.time(), '%I:%M %p'))`"
tz: 'America/New_York'
mainfont: Serif
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: false
    theme: cerulean
    highlight: zenburn
    highretina: 3
    self_contained: yes
    css: [../style.css]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  echo = FALSE,
  fig.align = 'center',
  warning = FALSE,
  message = FALSE,
  cache = FALSE,
  results='hide',
  fig.keep='all',
  fig.width = 7,   
  fig.height = 5,   
  out.width = "80%",
  dpi = 300         
)
```

```{r}
librarian::shelf(ggplot2, gt, readr)
  plots <- read_rds('../data/plots')
```



# Fiscal Impact Measure
```{r fiscal_impact}
plots$fiscal_impact_measure
```



# Summary 
```{r, include=TRUE, results = 'show'}
summary_tbl
```


# FIM Contribution
````{r fim_contribution}
plots$federal_contribution
plots$state_contribution
plots$consumption_contribution

````

# FIM Consistent Components
````{r fim_component_real}
plots$fim_federal_purchases_real
plots$fim_state_purchases_real
plots$consumption_real

````

# NIPA Consistent Components
````{r nipa_component_real}
plots$nipa_federal_purchases_real
plots$nipa_state_purchases_real

````






# Grants {.tabset .tabset-pills}

## Contributions
```{r grants}
plots$consumption_grants_contribution
plots$investment_grants_contribution
```

## Real Levels
```{r grants_rlevels}
plots$consumption_grants_real
plots$investment_grants_real
```

## Nominal Levels
```{r grants_levels}
plots$consumption_grants
plots$investment_grants
```

# Taxes

## Corporate Taxes {.tabset .tabset-pills}

### Contributions
```{r corp_contribution}
plots$federal_corporate_taxes_contribution
plots$state_corporate_taxes_contribution
```

### Real Levels
```{r corp_rlvl}
plots$federal_corporate_taxes_real
plots$state_corporate_taxes_real
```

### Nominal Levels
```{r corp_lvl}
plots$federal_corporate_taxes
plots$state_corporate_taxes
```

## Non-Corporate Taxes {.tabset .tabset-pills}

### Contributions
```{r noncorp_cont}
plots$federal_non_corporate_taxes_contribution
plots$state_non_corporate_taxes_contribution
```

### Real Levels
```{r noncorp_rlvl}
plots$federal_non_corporate_taxes_real
plots$state_non_corporate_taxes_real
```

### Nominal Levels
```{r noncorp_lvl}
plots$federal_non_corporate_taxes
plots$state_non_corporate_taxes
```

## Supply Side IRA {.tabset .tabset-pills}

### Contributions
```{r ira_cont}
plots$supply_side_ira_contribution
```

### Real Levels
```{r ira_rlvl}
plots$supply_side_ira_real
```

### Nominal Levels
```{r ira_lvl}
plots$supply_side_ira
```

# Transfers

## Total
```{r transfers_contribution}
plots$transfers_contribution
```

## Health Outlays {.tabset .tabset-pills}

### Contributions
```{r health}
plots$federal_health_outlays_contribution
plots$state_health_outlays_contribution
```

### Real Levels
```{r health_rlvl}
plots$federal_health_outlays_real
plots$state_health_outlays_real
```

### Nominal Levels
```{r health_lvl}
plots$federal_health_outlays
plots$state_health_outlays
```

## Subsidies (non-ARP) {.tabset .tabset-pills}

### Contributions
```{r subsidies}
plots$federal_subsidies_contribution
plots$state_subsidies_contribution
```

### Real Levels
```{r subsidies_rlvl}
plots$federal_subsidies_real
plots$state_subsidies_real
```

### Nominal Levels
```{r subsidies_lvl}
plots$federal_subsidies
plots$state_subsidies
```

## Subsidies (ARP) {.tabset .tabset-pills}
ARP PPP and Provider Relief Fund

### Contribution
```{r subsidies-arp-cont}
plots$federal_aid_to_small_businesses_arp_contribution
```

### Real Levels
```{r subsidies-arp-rlevel}
plots$federal_aid_to_small_businesses_arp_real
```

### Nominal Levels
```{r subsidies-arp-level}
plots$federal_aid_to_small_businesses_arp
```

## Unemployment Insurance {.tabset .tabset-pills}

### Contributions
```{r ui_cont}
plots$federal_ui_contribution
plots$state_ui_contribution
```

### Real Levels
```{r ui_rlvl}
plots$federal_ui_real
plots$state_ui_real
```

### Nominal Levels
```{r ui_lvl}
plots$federal_ui
plots$state_ui
```

## Other Aid to Vulnerable Households {.tabset .tabset-pills}
Premium Tax Credits, Ratepayer protection, Assistance for older Americans, COBRA, Emergency Assistance

### Contribution
```{r vulnerable}
plots$federal_other_vulnerable_arp_contribution
```

### Real Level
```{r vulnerable_rlvl}
plots$federal_other_vulnerable_arp_real
```

### Nominal Level
```{r vulnerable_lvl}
plots$federal_other_vulnerable_arp
```

## Rebate Checks (1st and 2nd round) {.tabset .tabset-pills}

### Contribution
```{r rebate_cont}
plots$rebate_checks_contribution
```

### Real Level
```{r rebate_rlvl}
plots$rebate_checks_real
```

### Nominal Level
```{r rebate_lvl}
plots$rebate_checks
```

## Rebate Checks (ARP) {.tabset .tabset-pills}

### Contribution
```{r rebate_arp_cont}
plots$rebate_checks_arp_contribution
```

### Real Level
```{r rebate_arp_rlvl}
plots$rebate_checks_arp_real
```

### Nominal Level
```{r rebate_arp_lvl}
plots$rebate_checks_arp
```

## Direct Aid to Households {.tabset .tabset-pills}
Includes Child tax credit, EITC, Childcare for workers, Dependent care for families

### Contribution
```{r aid_cont}
plots$federal_other_direct_aid_arp_contribution
```

### Real Level
```{r aid_rlvl}
plots$federal_other_direct_aid_arp_real
```

### Nominal Level
```{r aid_lvl}
plots$federal_other_direct_aid_arp
```

## Social Benefits Remainder {.tabset .tabset-pills}

### Contribution
```{r soc_cont}
plots$federal_social_benefits_contribution
plots$state_social_benefits_contribution
```

### Real Level
```{r soc_rlvl}
plots$federal_social_benefits_real
plots$state_social_benefits_real
```

### Nominal Level
```{r soc_lvl}
plots$federal_social_benefits
plots$state_social_benefits
```

## Student Loans {.tabset .tabset-pills}

### Contribution
```{r student_loans_cont}
plots$federal_student_loans_contribution
```

### Real Level
```{r student_loans_rlvl}
plots$federal_student_loans_real
```

### Nominal Level
```{r student_loans_lvl}
plots$federal_student_loans
```

