---
title: "FIM Technical Methodology"
author: "Lorae Stojanovic"
date: "`r Sys.Date()`"
output: 
  bookdown::gitbook:
    split_by: chapter
site: bookdown::bookdown_site
documentclass: book
# Bookdown is a required package to run this document. If you don't already
# have it installed, simply run:
# install.packages("bookdown")
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(DiagrammeR)
library(DT)

```

To serve this book, run the following code in the R console: (Sorry, I know the working directory stuff is trash. But there are too many extraneous Rmd files clogging the root directory, and until I delete them all, this is the best solution I have)
```{r}
# Save the current working directory. Must be the root folder of the project.
old_dir <- getwd()

# Change to the book directory
setwd("docs/technical_methodology")

# Render the book
bookdown::render_book("index.Rmd", output_dir = "_book")

# Revert to the original working directory
setwd(old_dir)

```

# Welcome!
The purpose of this document is to explain, in depth, how the FIM works. It is intended to encapsulate the technical elements of the project arising from the programming design as well as the economic design.

Chapter 1 documents the sources of data.

Chapter 2 documents how the FIM is calculated.

Chapter 3 documents how the UI interface is constructed.

Chapter 4 describes the economic intuition for the FIM.

```{r flow_chart, echo=FALSE}
grViz("
digraph flowchart {
  node [fontname = Helvetica, shape = box]

  Start [label = 'Start']
  Data_Input [label = 'Data Input']
  Data_Cleaning [label = 'Data Cleaning']
  Analysis [label = 'Analysis']
  Report_Generation [label = 'Report Generation']
  End [label = 'End']

  Start -> Data_Input
  Data_Input -> Data_Cleaning
  Data_Cleaning -> Analysis
  Analysis -> Report_Generation
  Report_Generation -> End
}
")
```
